---
title: "pre_lab_01.Rmd"
author: "Derek Willis"
date: "2023-01-05"
output: html_document
---

```{r setup, include=FALSE}
install.packages('knitr')

knitr::opts_chunk$set(echo = TRUE)
options(scipen=999)
```

## About this notebook

This notebook contains code and explanatory text that your should review and run as you read through chapters 13 and 15 of the course textbook, "Data Journalism with R and the Tidyverse". Answer questions and edit the document as directed.

Running this notebook will help you understand key data analysis methods and concepts that you will put into practice during this week's lab. When you come across a new command or verb in R, you should add it to your class reference notebook file along with what it's used for.

When you are finished running the code in this notebook, you will push changes to your course GitHub repo, and upload the link to ELMS as instructed.

You also should update your class reference notebook with code examples and what they mean. When you come across a new concept or R verb, put it in your reference notebook along with a sentence about what it does. For example:

```{r}
install.packages('tidyverse')
library(tidyverse)
```

The library function loads an R library such as the tidyverse, providing access to the verbs and functions that library contains. You have to load a library before you can use its functions.

## Instructor in class demonstration outline

-   Load this notebook, and provide overview of R, RStudio (text editor, console, environment window, file explorer), Rmarkdown data notebooks, the reference notebook.
-   Work through the code below, chapter by chapter.
    -   Chapter 13: explain the console and the concept of variables. Explain packages, the Tidyverse, how to install them.
    -   Chapter 15: explain loading libraries, the concept of functions, loading data, the importance of having questions to guide data analysis, and aggregation and summarization.

## Chapter 13

### Task 1: The console

**Task** In the console (look below), type 2+2 and hit enter. What does it return? Type your answer below. **Answer:**

### Task 2: Variables

**Task** Put the following code -- which stores the numeral 2 as a variable called number -- in the console and hit enter.

```{r}
number <- 2
```

Then type this code to multiply the variable called number by itself.

```{r}
number * number
```

What does it return? Type your answer below:

**Answer:**

### Task 3: Install packages

**Task** In the console, run the following code to install the Tidyverse:

```{r}
install.packages('tidyverse')
```

**Task** Add that

Then run the following code to install the rmarkdown library

```{r}
install.packages('rmarkdown')
```

### Task 4: Load libraries

**Task** Run the following code in the gray-colored codeblock below -- not in the console -- to load the tidyverse library. To run the code, click the little green play button (left facing arrow) at the top right of the codeblock. In Rmarkdown data notebooks, we write code inside of codeblocks, and explanatory text in the white area outside of it.

```{r}
# This is a comment inside of a codeblock. Comments doesn't run when we run the codeblock.
# Adding this command is a good way to start your class reference notebook.
library(tidyverse)
```

### Task 5: Load data

**Task** Load the UMD courses data by running the following codeblock. The data is contained in an .rds file, which is a binary format for R data.

```{r}
umd_courses <- read_rds("umd_courses.rds")
```

Once it loads, look at the environment window at right in RStudio for umd_courses It's next to a small blue circle with a white arrow in it. To the right of that, look at how many "obs" the dataset has ("obs" is short for observation, which is another way of saying rows or records). Look at how many variables it has (another way of saying columns or fields). Write the number of observations and variables below.

**Answer**

### Task 6: Glimpse data

**Task** Run the following codeblock to use glimpse() to get a sense of the data, some of the values in each field/variable/column, and the data types of each field/variable/column.

```{r}
glimpse(umd_courses)
```

In the space below, write the data type for the "id" column and the "seats" column (it's in light gray, offset by \< \>). **Answer**

### Task 7: Examine data "head"

**Task** Run the following codeblock to use head() to examine the first six rows of data. Explore the data by clicking the black arrow on the right side of the output table, allowing you to examine additional columns.

```{r}
head(umd_courses)
```

In the space below, write the name of the business from the first row of the "name" column. **Answer**

### Task 8: Explore data with interactive dataframe

**Task** look at the environment window at right in RStudio for umd_courses. It's next to a small blue circle with a white arrow in it. Click on the umd_courses -- not the blue circle. It will open the dataframe as an interactive table. Scroll down to see additional rows. Scroll left to see additional columns. Click on a column name to sort by that column.

### Task 9: Group and count

So what if we wanted to know how many classes were offered in each term?

To do that by hand, we'd have to take each of the 60,672 individual rows (or observations or records) and sort them into a pile. We'd put them in groups -- one for each term -- and then count them. To do this, we'll take our dataset and we'll introduce a new operator: `|>`. The best way to read that operator, in my opinion, is to interpret that as "and then do this." The keyboard shortcut for this is cmd-shift-m on the Mac and ctrl-shift-m on Windows. Don't like that character? R now has a new one that does the same thing: `|>`.

**Task** Run the following codeblock:

```{r}
umd_courses |>
  group_by(term) |>
  summarise(
    count_classes = n()
  )
```

Describe what you think each line of this code is doing below, based on the result:

**Answer**

### Task 10: Adding arrange to sort the data

**Task** Run the following codeblock:

```{r}
umd_courses |>
  group_by(term) |>
  summarise(
    count_classes = n()
  ) |>
  arrange(desc(count_classes))
```

Compared to the previous codeblock, what has changed and how did that affect the result? **Answer**

### Task 11: Grouping by multiple columns

**Task** Run the following codeblock.

```{r}
umd_courses |>
  group_by(term, department) |>
  summarise(
    count_classes = n()
  ) |>
  arrange(term, department)
```

Compared to the previous codeblock, how did the output change and what in the code changed it?

**Answer**

### Task 12: Other summary functions

**Task** Run the following codeblock:

```{r}
umd_courses |>
  group_by(term) |>
  summarise(
    count_classes = n(),
    total_seats = sum(seats)
  ) |>
  arrange(desc(total_seats))
```

Compared to the previous codeblock, what has changed and how did that affect the result? **Answer**

### Task 13: different kinds of summarizations

**Task** Run the following codeblock to group the umd_courses dataframe by term. Then do the following summarizations for each term: count the number of courses, add up the total of all seats, and calculate the mean and median contribution number of seats. Then sort so the term with the most total seats is listed first.

```{r}
umd_courses |>
  group_by(term) |>
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats = mean(seats),
    median_seats = median(seats)
  ) |>
  arrange(desc(total_seats))
```

Compared to the previous codeblock, what has changed and how did that affect the result? What do the top terms have in common? **Answer**

### Task 14: more aggregations

**Task** Run the following codeblock to group the umd_courses dataframe by the department. Then do the following summarizations for each group: count the number of courses, add up the total of all seats, calculate the mean and median number of seats and then add lines to calculate the smallest and largest seat numbers using the `min()` and `max()` functions. Then sort so the department with the largest total number of seats is listed first.

```{r}
umd_courses |>
  group_by(department) |>
  summarise(
    count_classes = n(),
    total_seats = sum(seats),
    mean_seats = mean(seats),
    median_seats = median(seats),
    min_seats = min (seats),
    max_seats = max(seats)
    # add minimum seats here
    # add maximum seats here
  ) |>
  arrange(desc(total_seats))
```

Compared to the previous codeblock, what has changed and how did that affect the result? How is the data sorted?

**Answer**

### Task 15: Questions

No code here; I want you to provide three questions that this data suggests to you. What would you like to ask it?

**Answer**

1.  First question I'd want to ask is, what is the number of seats by class for each professor within any department teaching the same course? Are there any disparities in number of available seats? I'd then want to figure how how to interpret that. For example, does X professor have many available seats because of popular demand, or is it because the professor isn't popular among students? Does Y professor have few available seats because there's such a high demand for his/her/their classes, such that it's overwhelming and department officials have to limit seat numbers in some way? Or, could so few seats mean the opposite is true: Is there such little demand for this professor that the department isn't expecting many students in those classes, and is thus making only a few seats open? From there, as a data journalist, my inclination is to first seek aggregate data like average number of stars in ratemyprofessor.com ratings from students. To me, If I'm seeing a glaring disparity in these ratings, too, then I feel confident in the strength and significance of the correlation I find, whichever the correlation ends up being.

2.  Another question that comes to mind is, how significantly do available seats vary among both elective and required courses in any given department? A trend I'd expect would be that required courses would have more available seats in order to accommodate enough people so they may satisfy critical requirements, and that less important electives would have fewer seats because the nature of them not being required wouldn't elicit high demand for students as they're scheduling. One factor I'd keep in mind is that there could reasonably likely be a trend that those required classes are in much larger venues like lecture halls in order to accommodate more students, and that electives might generally be in smaller classrooms. This aforementioned thing for me to keep in mind would suggest to me that the answer to my question/prediction is obvious–that core courses are for classes with many seats available and that electives have fewer seats–but it would also be a smart and simple metric by which I determine whether there's the correlation that I believed there would be.

3.  Another question I have is, comparing among departments, which of them have significantly many–or significantly few–classes offered in any given semester? That question probably sounds odd, but what I'm thinking about in relation to this is, given that there are various capstone projects across colleges at UMD, it's reasonable to suspect that not a particularly large number of courses would be offered while students are enrolled in capstones, which I understand are often more rigorous and demanding. For me, as I understand it, I'll be in such a program for my final semester here and that I likely don't really have any other classes to take at the same time as it. So, cutting to the chase, I'm curious as to whether any colleges here in particular have capstones that are rigorous to the extent that students involved don't need to take any classes at the same time. Conversely, I wonder whether some programs that aren't particularly demanding in comparison offer more classes as a result. The obvious first step is to first analyze data regarding number of classes offered per semester for each department; in doing this, I'd ascertain whether there's any disparity/distinction of interest in the first place. Then, if I were right that there is such an observation to to be made, I would proceed to investigate each unit's respective capstone program, making note of what variables I could use to make comparisons in difficulty/prestige by college/department. I'd want to see what each grading scale requires, perhaps. Or, maybe I'd see if they differ in number of credits or amount of time students are required to dedicate toward completion/satisfying requirements. Maybe even the metric of whether it's individual or group work could bear relevance in determining if a program is easier or harder.
