---
title: "lab_01"
author: "Derek Willis"
adapted by: "Daniel Trielli"
date: "2024-05-23"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# Turn off scientific notation
options(scipen=999)
```

## About this lab

To complete this lab, you need to: \* run existing code as directed (look for **Task**). \* modify existing code as directed (look for **Task**). \* write code in empty codeblocks provided to answer questions included (look for **Q**). \* write out the answer in the form of a complete sentence in the space given (look for **A**). When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

Remember to add any new verbs or concepts to your class reference notebook!

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Load the tidyverse. If you have not installed the tidyverse already, remove the # from the next line and run it first.  
# install.packages('tidyverse')
library(tidyverse)
```

## Load Data

We're going to load the UMD courses data.

**Task** Run the codeblock below to load the data, which is contained in an RDS file.

```{r}
umd_courses <- read_rds("lab_01.rds")
```

## Answer questions

### **Q1** Which department has the highest mean (average) number of seats for its classes? You need to show the department with the highest average first. Which one is the most surprising/interesting to you?

mean (artstudio) = 

mean (musicperformance) =

mean (immigration studies) =

mean (anthropology) =

mean (philosophy) = 

mean (scholarssocietysciencetechnology) =

mean (film studies) =

mean (neuroscience&cognitivescience) = 

mean (biometrics) = 

mean (scholarsarts) =

mean (scholarslifesciences)

mean (classics) =

mean (reliabilityengineering) = 

mean (theatre) =

mean (germanicstudies) = 

mean (higheredstudentaffairs&internationaledpolicy) =

mean (historic preservation) =

mean (korean) =

mean (nutrition&foodsciences) =

mean (bioengineering) = 

mean (schoolofmusic) = 

mean (scholars&internationalstudies) =

mean (publicpolicy) = 

mean (science&globalchange) = 

mean (telecommunications) = 

mean (biology) =

mean (kinesiology) =

mean (marineesturarineenvironmentalsciences) = 

mean (landscapearchitecture) = 

mean (computerscience) = 

I've played around with the different ways to enter these, including no spaces and using or not using and or &, but I'm still not getting it to calculate results. Visually, my educated guess would be that the answer is school of music because it appears it has the most slots on the spreadsheet. 

```{r}

```

**A1**

------------------------------------------------------------------------

### **Q2** Find out the greatest number of times any individual class has been offered in our dataset. You must show the greatest number of classes first.

```{r}

```

**A2** Still having trouble getting the formular entered correctly, but my guess, visually, is still school of music based on what appears to have most boxes. 

------------------------------------------------------------------------

### **Q3** Count the number of courses for each instructor and department. Which instructor has taught the most?

```{r}

```

**A3** I'm taking another best guess despite not having success getting the formulas entered properly, but I believe it's someone within school of music. 

------------------------------------------------------------------------

### **Q4** Given the results for Q3, what would have made it easier to find the answer? You don't have to know *how* to accomplish this; I want you to describe in words what would make the answer easier to see.

```{r}

```

**A4**

------------------------------------------------------------------------

### **Q5** Take one of the questions you listed in the final question of the pre_lab and try to answer it here in code. If you don't know how to fully answer it, write what code you can and describe in words what else you would need to do. If you got an answer, is it interesting?

```{r}

```

**A5** I'm working on accessing my completed pre-lab, as I think I didn't save it properly and that it might've been deleted by mistake.
